humanCdGF_predict_cfu

### 2016-10-10 code/HumanCdGF_predict_cfu.Rmd
add median CFU prediction

heat.plot <- function(input_maxtrix, high_color = '#660000', low_color = '#FFFFFF', legend_title = NA) {
  long.df <- data.frame(input_maxtrix)
  colnames(long.df) <- c(1:10)
  long.df$community_day <- as.character(c(0:10))
  long.df %>% gather(cfu_day, variable, -community_day) %>% 
    ggplot(aes(cfu_day, community_day)) + 
    geom_tile(aes(fill = variable), color = 'white') +
    scale_fill_gradient(low = low_color, high = high_color, name = legend_title) +
    labs(x = 'Day (CFU)', y = 'Day (Community)') + 
    theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank()) + 
    scale_x_discrete(expand = c(0, 0), limits = c('1','2','3','4','5','6','7','8','9','10')) + 
    scale_y_discrete(expand = c(0, 0), limits = rev(c('0','1','2','3','4','5','6','7','8','9','10'))) + 
    theme(legend.position = 'top', legend.key.width = unit(2, 'cm'))
}
```

@@

Removing these data points improves predictability for day 6 and 8 but not the earlier days. If I increase the cutoff to remove all CFU below 4, there is still no improvement and even in some cases decreased performance.

@@

>>>>

heat.plot <- function(input_maxtrix, high_color = '#660000', low_color = '#FFFFFF', legend_title = NA) {
  long.df <- data.frame(input_maxtrix)
  colnames(long.df) <- c(1:10)
  long.df$community_day <- as.character(c(0:10))
  long.df %>% gather(cfu_day, variable, -community_day) %>% 
    ggplot(aes(cfu_day, community_day)) + 
    geom_tile(aes(fill = variable), color = 'white') +
    scale_fill_gradient(low = low_color, high = high_color, name = legend_title, limits = c(0,1)) +
    labs(x = 'Day (CFU)', y = 'Day (Community)') + 
    theme(axis.line.x = element_blank(), axis.line.y = element_blank(), axis.ticks = element_blank()) + 
    scale_x_discrete(expand = c(0, 0), limits = c('1','2','3','4','5','6','7','8','9','10')) + 
    scale_y_discrete(expand = c(0, 0), limits = rev(c('0','1','2','3','4','5','6','7','8','9','10'))) + 
    theme(legend.position = 'top', legend.key.width = unit(2, 'cm'))
}
```

@@

Removing these data points improves predictability for day 6 and 8 but not the earlier days. If I increase the cutoff to remove all CFU below 5, there is a slight increase in day 3 cfu classification at the cost of decrease day 1 cfu classification.

@@

### Does taking the median CFU improve predicatbility?

```{r predict median CFU, fig.width=6,fig.height=6}

rsq_med <- matrix(NA, nrow = 11, ncol = 10)
otu_votes_med <- rep(0, ncol(rel_abund_shared_01))
names(otu_votes_med) <- colnames(rel_abund_shared_01)
for(n in 1:10){
med_log_cfu <- meta_data %>% 
  filter(day %in% c(n:10), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  select(mouse_id, cage_id, log_cfu) %>% 
  group_by(mouse_id, cage_id) %>% 
  summarise(log_cfu_med = median(log_cfu)) %>% ungroup()
for(m in 0:10){
  sample_list <- shared_list_days %>% 
    filter(day == m) %>% 
    select(-day)
  community.df <- rel_abund_shared_01 %>% 
    right_join(sample_list) %>% 
    select(-sample_id)
  cfu_shared.df <- med_log_cfu %>% 
      select(mouse_id, log_cfu_med) %>% 
      right_join(community.df) %>% 
      select(-mouse_id) %>% 
#      select(log_cfu_med, one_of(top_n_overall)) %>% 
      filter(!is.na(log_cfu_med))
    set.seed(seed)
    rf_top <- randomForest(log_cfu_med ~ ., data = cfu_shared.df,
                           importance = TRUE, ntree = n_trees)
    importance_sorted <- sort(importance(rf_top)[,1], decreasing = T)
    top_important_OTU <- rownames(data.frame(head(importance_sorted, n_features)))
    set.seed(seed)
    rf_n_features <- randomForest(log_cfu_med ~ ., data=cfu_shared.df[ ,c( 'log_cfu_med',
                                                                       names(importance_sorted)[1:n_features])], 
                                  importance=TRUE, ntree=n_trees)
    importance_final <- sort(importance(rf_n_features)[,1], decreasing = T)
    top_OTU_final <- rownames(data.frame(head(importance_final, n_features)))
    rsq_med[m+1, n] <- round(tail(rf_n_features$rsq, 1), 2)
    otu_votes_med <- otu_votes_med + 1*(colnames(rel_abund_shared_01) %in% top_OTU_final)
}
}

heat.plot(rsq_med, '#003300', legend_title = expression(paste('R'^'2'))) + 
  ggtitle('Day CFU is median value of Day n to Day 10')

```
